The Impact of Interview Disruptions on Data Quality in Social Surveys

### 2024-12-03 paper/paper.qmd
update paper

In recent years, the quality and reliability of data collected through social surveys have garnered increasing attention, especially considering how various forms of disruption may influence respondents’ behavior and, subsequently, the quality of survey responses. Social surveys are indispensable tools for obtaining observations into societal trends, guiding public policy, and informing evidence-based decision-making. Thus, ensuring the robustness of data collected from these surveys has implications for the accuracy of the conclusions drawn and for the effective formulation of policy measures. Previous research has made significant strides in understanding factors that impact survey data quality. For example, Groves et al. [@groves2011survey] investigated the influence of interviewer characteristics—such as demeanor, experience, and communication style—on the quality of data obtained, employing a multi-level modeling approach to assess these refined dynamics. Similarly, Couper and de Leeuw [@massey2013we] explored how the survey mode (e.g., in-person versus telephone) affects respondent comfort and attentiveness, highlighting that face-to-face interviews often elicit more thorough responses due to the availability of visual cues. In another significant contribution, Fowler and Mangione [@aquilino1993effects] analyzed the impact of third-party presence during interviews, finding that interruptions can lead to response biases, thereby compromising data integrity. Despite these advances, there remains a clear gap that needs to be filled: the distinct effects of different types of disruptions—specifically those involving family members or occurring due to environmental factors—on respondent comprehension and the overall quality of responses have not yet been systematically explored. Addressing this gap is paramount to improving the reliability of survey data, particularly in settings where such disruptions are unavoidable.

@@

## Overview
We utilized the statistical programming language R [@citeR] to prepare, analyze, and visualize the survey data. This study focuses on understanding how interview disruptions affect response quality by examining variables related to both interviewer and respondent characteristics. The structured design of the dataset allowed for a thorough exploration of contextual disruptions and their influence on data reliability, providing valuable insights for enhancing survey methodologies.

@@

- **arrow** [@arrow]: Employed for efficiently reading and writing large datasets, enhancing data handling across different systems.

- **knitr** [@knitr]: Used to create dynamic documents with R, including reports and presentations. It enables seamless integration of R code and text.

- **rstanarm** [@rstanarm]: Provides tools for Bayesian regression modeling using the 'Stan' platform, making it straightforward to apply Bayesian models in a regression context.

@@

In addition to age, we included variables such as "interview conduct method," various "interference types" (e.g., "partner_interference," "child_interference"), and "interviewer characteristics" to fully encapsulate the different factors known to impact response quality. These variables were selected based on their relevance in literature, including the work of Fowler and Mangione [@aquilino1993effects], which underscores the importance of interviewer attributes and environmental conditions in shaping survey data reliability. By including such factors, our model was able to reflect a realistic and integrated picture of the dynamics at play during social survey interviews, thereby providing a robust framework for analyzing how disruptions influence response accuracy.

## Alternative model
After evaluating both the decision tree model [@tbl-comprehension-matrix13] and the Bayesian GLM [@tbl-comprehension-matrix12] for classifying comprehension outcomes, a clear distinction in performance emerged. While the decision tree model demonstrated superior sensitivity and balanced accuracy in identifying the minority class, "Not_understood," the Bayesian GLM consistently achieved higher overall accuracy, with a score of 96.05% compared to the decision tree's 95.98%.

@@

# Confusion matrix definition
confusion_matrix <- matrix(c(41, 36, 129, 3970), nrow = 2, byrow = TRUE,
                           dimnames = list("Prediction" = c("Not_understood", "Understood"),
                                           "Reference" = c("Not_understood", "Understood")))

@@

-   $\beta_{12}$: The coefficient for **respondent tried_best**, indicating whether the respondent tried their best to provide accurate answers during the survey. Categories include:

@@

The model is specified as follows for ordered logistic regression:\begin{align}
y_i &\sim \text{OrderedLogistic}(\eta_i, \kappa) \\
\eta_i &= \beta_0 + \beta_{\text{interview\_conduct\_method}} \times \text{interview\_conduct\_method}_i \notag \\
       &\quad + \beta_{\text{partner\_interference}} \times \text{partner\_interference}_i \notag \\
       &\quad + \beta_{\text{child\_interference}} \times \text{child\_interference}_i 
       + \beta_{\text{parent\_interference}} \times \text{parent\_interference}_i \notag \\
       &\quad + \beta_{\text{non\_relative\_interference}} \times \text{non\_relative\_interference}_i 
       + \beta_{\text{interviewer\_age}} \times \text{interviewer\_age}_i \notag \\
       &\quad + \beta_{\text{interviewer\_gender}} \times \text{interviewer\_gender}_i 
       + \beta_{\text{country}} \times \text{country}_i \notag \\
       &\quad + \beta_{\text{question\_clarification}} \times \text{question\_clarification}_i 
       + \beta_{\text{respondent\_reluctant}} \times \text{respondent\_reluctant}_i \notag \\
       &\quad + \beta_{\text{respondent\_tried\_best}} \times \text{respondent\_tried\_best}_i \\
\beta &\sim \text{Normal}(0, 10) \quad (\text{default non-informative prior}) \\
\kappa &\sim \text{Normal}(0, 5) \quad (\text{default prior for cutpoints})
\end{align}

@@

# Predict probabilities and convert predicted_class to "Understood" and "Not_understood"
predicted_prob <- predict(first_model, newdata = analysis_data_test, type = "response")
analysis_data_test$predicted_class <- ifelse(predicted_prob > 0.5, "Understood", "Not_understood")

# Convert predicted_class and respondent_understood_binary to factors with consistent levels
analysis_data_test$predicted_class <- factor(analysis_data_test$predicted_class, levels = c("Not_understood", "Understood"))
analysis_data_test$respondent_understood_binary <- factor(analysis_data_test$respondent_understood_binary, levels = c("Not_understood", "Understood"))

@@

confusion_matrix <- matrix(c(41, 36, 129, 3970), nrow = 2, byrow = TRUE,
                           dimnames = list("Prediction" = c("Not_understood", "Understood"),
                                           "Reference" = c("Not_understood", "Understood")))

# Extract True Positive, False Positive, True Negative, and False Negative from the confusion matrix
TP <- confusion_matrix["Not_understood", "Not_understood"]
FP <- confusion_matrix["Not_understood", "Understood"]
TN <- confusion_matrix["Understood", "Understood"]
FN <- confusion_matrix["Understood", "Not_understood"]

@@

```
@tbl-comprehension-matrix1 provides a summary of the model's prediction outcomes for the comprehension classification task. The table displays the frequencies of correct and incorrect predictions across the "Not_understood" and "Understood" categories. Correct classifications occur when the predicted and actual labels align (e.g., "Not_understood" predicted as "Not_understood"), while mismatches are shown as incorrect classifications (e.g., "Not_understood" predicted as "Understood"). This representation effectively highlights the model's ability to distinguish between the two classes, particularly its challenges in identifying the minority "Not_understood" category due to its smaller sample size.

@tbl-comprehension-matrix2 presents key performance metrics for evaluating the classification model. Metrics such as Accuracy, Precision (Positive Predictive Value), Recall (True Positive Rate), Specificity (True Negative Rate), F1 Score, and Balanced Accuracy provide a detailed view of the model's strengths and limitations. The high Accuracy and Specificity values demonstrate the model's reliability in correctly classifying the majority "Understood" category. In contrast, the lower Recall for the "Not_understood" class indicates difficulties in detecting this minority group. These metrics offer a well-rounded understanding of the model's performance, balancing its effectiveness in majority class classification with areas for improvement in minority class sensitivity.

@@

The significance of interview methods also warrants attention. Notably, the positive influence observed for "interview_conduct_method3" suggests that certain approaches to interviewing—potentially those incorporating more interaction or structured guidance—can substantially enhance comprehension. Training interviewers in these methods, which may include explicit clarity and engagement-promoting techniques, could lead to more consistent and high-quality data, particularly in settings prone to interference.

@@

The overarching lesson from this study is the significance of managing interview conditions to optimize respondent comprehension and, subsequently, data quality. Minimizing household disruptions, particularly familial interference, appears to be a essential factor in ensuring the accuracy and reliability of survey responses. Pre-interview guidelines that advise respondents on how to create a distraction-free environment could help mitigate these issues. Furthermore, adopting interactive and structured interview techniques, akin to the positively influential "interview_conduct_method3," could foster greater respondent engagement and understanding.

@@

To enhance clarity and usability, several column names were renamed to be more descriptive. For example, resundq was renamed to respondent_understood. Additionally, the respondent_understood variable was recoded into a binary form (Understood vs. Not_understood) to simplify future analyses.

>>>>

In recent years, the quality and reliability of data collected through social surveys have garnered increasing attention, especially considering how various forms of disruption may influence respondents’ behavior and, subsequently, the quality of survey responses. Social surveys are indispensable tools for obtaining observations into societal trends, guiding public policy, and informing evidence-based decision-making. Thus, ensuring the robustness of data collected from these surveys has implications for the accuracy of the conclusions drawn and for the effective formulation of policy measures. Previous research has made significant strides in understanding factors that impact survey data quality. For example, Groves et al. [@groves2011survey].investigated the influence of interviewer characteristics—such as demeanor, experience, and communication style—on the quality of data obtained, employing a multi-level modeling approach to assess these refined dynamics. Similarly, Couper and de Leeuw [@massey2013we].explored how the survey mode (e.g., in-person versus telephone) affects respondent comfort and attentiveness, highlighting that face-to-face interviews often elicit more thorough responses due to the availability of visual cues. In another significant contribution, Fowler and Mangione [@aquilino1993effects]. analyzed the impact of third-party presence during interviews, finding that interruptions can lead to response biases, thereby compromising data integrity. Despite these advances, there remains a clear gap that needs to be filled: the distinct effects of different types of disruptions—specifically those involving family members or occurring due to environmental factors—on respondent comprehension and the overall quality of responses have not yet been systematically explored. Addressing this gap is paramount to improving the reliability of survey data, particularly in settings where such disruptions are unavoidable.

@@

## Overview
We utilized the statistical programming language R [@citeR]. to prepare, analyze, and visualize the survey data. This study focuses on understanding how interview disruptions affect response quality by examining variables related to both interviewer and respondent characteristics. The structured design of the dataset allowed for a thorough exploration of contextual disruptions and their influence on data reliability, providing valuable insights for enhancing survey methodologies.

@@

- **arrow** [@arrow].: Employed for efficiently reading and writing large datasets, enhancing data handling across different systems.

- **knitr** [@knitr].: Used to create dynamic documents with R, including reports and presentations. It enables seamless integration of R code and text.

- **rstanarm** [@rstanarm].: Provides tools for Bayesian regression modeling using the 'Stan' platform, making it straightforward to apply Bayesian models in a regression context.

@@

In addition to age, we included variables such as "interview conduct method," various "interference types" (e.g., "partner interference," "child interference"), and "interviewer characteristics" to fully encapsulate the different factors known to impact response quality. These variables were selected based on their relevance in literature, including the work of Fowler and Mangione [@aquilino1993effects], which underscores the importance of interviewer attributes and environmental conditions in shaping survey data reliability. By including such factors, our model was able to reflect a realistic and integrated picture of the dynamics at play during social survey interviews, thereby providing a robust framework for analyzing how disruptions influence response accuracy.

## Alternative model
After evaluating both the decision tree model [@tbl-comprehension-matrix13] and the Bayesian GLM [@tbl-comprehension-matrix12] for classifying comprehension outcomes, a clear distinction in performance emerged. While the decision tree model demonstrated superior sensitivity and balanced accuracy in identifying the minority class, "Not understood," the Bayesian GLM consistently achieved higher overall accuracy, with a score of 96.05% compared to the decision tree's 95.98%.

@@

# Confusion matrix definition
confusion_matrix <- matrix(c(41, 36, 129, 3970), nrow = 2, byrow = TRUE,
                           dimnames = list("Prediction" = c("Not understood", "Understood"),
                                           "Reference" = c("Not understood", "Understood")))

@@

-   $\beta_{12}$: The coefficient for **respondent tried best**, indicating whether the respondent tried their best to provide accurate answers during the survey. Categories include:

@@

The model is specified as follows for ordered logistic regression:\begin{align}
y_i &\sim \text{OrderedLogistic}(\eta_i, \kappa) \\
\eta_i &= \beta_0 + \beta_{\text{interview\ conduct\ method}} \times \text{interview\ conduct\ method}_i \notag \\
       &\quad + \beta_{\text{partner\ interference}} \times \text{partner\ interference}_i \notag \\
       &\quad + \beta_{\text{child\ interference}} \times \text{child\ interference}_i 
       + \beta_{\text{parent\ interference}} \times \text{parent\ interference}_i \notag \\
       &\quad + \beta_{\text{non\ relative\ interference}} \times \text{non\ relative\ interference}_i 
       + \beta_{\text{interviewer\ age}} \times \text{interviewer\ age}_i \notag \\
       &\quad + \beta_{\text{interviewer\ gender}} \times \text{interviewer\ gender}_i 
       + \beta_{\text{country}} \times \text{country}_i \notag \\
       &\quad + \beta_{\text{question\ clarification}} \times \text{question\ clarification}_i 
       + \beta_{\text{respondent\ reluctant}} \times \text{respondent\ reluctant}_i \notag \\
       &\quad + \beta_{\text{respondent\ tried\ best}} \times \text{respondent\ tried\ best}_i \\
\beta &\sim \text{Normal}(0, 10) \quad (\text{default non-informative prior}) \\
\kappa &\sim \text{Normal}(0, 5) \quad (\text{default prior for cutpoints})
\end{align}

@@

# Predict probabilities and convert predicted_class to "Understood" and "Not_understood"
predicted_prob <- predict(first_model, newdata = analysis_data_test, type = "response")
analysis_data_test$predicted_class <- ifelse(predicted_prob > 0.5, "Understood", "Not understood")

# Convert predicted_class and respondent_understood_binary to factors with consistent levels
analysis_data_test$predicted_class <- factor(analysis_data_test$predicted_class, levels = c("Not understood", "Understood"))
analysis_data_test$respondent_understood_binary <- factor(analysis_data_test$respondent_understood_binary, levels = c("Not understood", "Understood"))

@@

confusion_matrix <- matrix(c(41, 36, 129, 3970), nrow = 2, byrow = TRUE,
                           dimnames = list("Prediction" = c("Not understood", "Understood"),
                                           "Reference" = c("Not understood", "Understood")))

# Extract True Positive, False Positive, True Negative, and False Negative from the confusion matrix
TP <- confusion_matrix["Not understood", "Not understood"]
FP <- confusion_matrix["Not understood", "Understood"]
TN <- confusion_matrix["Understood", "Understood"]
FN <- confusion_matrix["Understood", "Not understood"]

@@

```
@tbl-comprehension-matrix1 provides a summary of the model's prediction outcomes for the comprehension classification task. The table displays the frequencies of correct and incorrect predictions across the "Not understood" and "Understood" categories. Correct classifications occur when the predicted and actual labels align (e.g., "Not understood" predicted as "Not understood"), while mismatches are shown as incorrect classifications (e.g., "Not understood" predicted as "Understood"). This representation effectively highlights the model's ability to distinguish between the two classes, particularly its challenges in identifying the minority "Not understood" category due to its smaller sample size.

@tbl-comprehension-matrix2 presents key performance metrics for evaluating the classification model. Metrics such as Accuracy, Precision (Positive Predictive Value), Recall (True Positive Rate), Specificity (True Negative Rate), F1 Score, and Balanced Accuracy provide a detailed view of the model's strengths and limitations. The high Accuracy and Specificity values demonstrate the model's reliability in correctly classifying the majority "Understood" category. In contrast, the lower Recall for the "Not understood" class indicates difficulties in detecting this minority group. These metrics offer a well-rounded understanding of the model's performance, balancing its effectiveness in majority class classification with areas for improvement in minority class sensitivity.

@@

The significance of interview methods also warrants attention. Notably, the positive influence observed for "interview conduct method3" suggests that certain approaches to interviewing—potentially those incorporating more interaction or structured guidance—can substantially enhance comprehension. Training interviewers in these methods, which may include explicit clarity and engagement-promoting techniques, could lead to more consistent and high-quality data, particularly in settings prone to interference.

@@

The overarching lesson from this study is the significance of managing interview conditions to optimize respondent comprehension and, subsequently, data quality. Minimizing household disruptions, particularly familial interference, appears to be a essential factor in ensuring the accuracy and reliability of survey responses. Pre-interview guidelines that advise respondents on how to create a distraction-free environment could help mitigate these issues. Furthermore, adopting interactive and structured interview techniques, akin to the positively influential "interview conduct method3," could foster greater respondent engagement and understanding.

@@

To enhance clarity and usability, several column names were renamed to be more descriptive. For example, resundq was renamed to respondent understood. Additionally, the respondent understood variable was recoded into a binary form (Understood vs. Not understood) to simplify future analyses.
